---
title: "ANX_proximitySNP_plot"
output: html_document
date: "2026-06-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Proposition: we take “clusters” (of chromosomally close SNPs) and graph their p-values in these clusters
- and identify + label the most influential SNP in each one
OR: we could find what gene they are on and group them that way 
OR: both! we group clusters of SNPs that are close to each other (we can talk about what close means in
comments, etc. and cite the research we use) and identify what gene each cluster lies on. We can single out the most influential SNP of each cluster and label it

```{r}
library(ggplot2)
library(tidyverse)
library(tidyr)
library('readxl')

Xanax <- read_excel('ANX_2026_daner_fullANX_v12_woUTAH_11022026.gz.p4.clump.areator.sorted.1mhc (1).xls')

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
