---
title: "ANX_proximitySNP_plot"
output: html_document
date: "2026-06-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Proposition: we take “clusters” (of chromosomally close SNPs) and graph their p-values in these clusters
- and identify + label the most influential SNP in each one
OR: we could find what gene they are on and group them that way 
OR: both! we group clusters of SNPs that are close to each other (we can talk about what close means in
comments, etc. and cite the research we use) and identify what gene each cluster lies on. We can single out the most influential SNP of each cluster and label it

```{r}
library(ggplot2)
library(tidyverse)
library(tidyr)
library('readxl')

Xanax <- read_excel('ANX_2026_daner_fullANX_v12_woUTAH_11022026.gz.p4.clump.areator.sorted.1mhc (1).xls')

```

## TO DO: FOR THURSDAY
Step 1: have 2 data frames: 1 for chromosome positions 1-10, 11-22 (complete)
Step 2: subset data frames by "VERY CLOSE" (I think it's the 20k bp apart or under?) - Julia
Step 2.5: format code for screenshots - Jackie
Step 3: find the genes on which these "VERY CLOSE" SNPs lie on (and cite references) - Soumya, Ananya, Eva
Step 4: graph the "very close" SNPs on a bar-chart esque graph, separated by gene (one graph = one chromosome, there should be 22 graphs) - Soumya, Ananya, Eva
Step 5: shove manhattan and volcano into presentation and allot scripts/lines - Julia and Jackie


```{r pressure, echo=FALSE}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
