---
title: "ANX_volcano_plot"
output: html_document
Author: Soumya Kalle
---

```{r}
print("Hello, world!")
```

Step 1: Load Anxiety GWAS dataset
```{r}
install.packages("readxl")
library('readxl')

ANX_dataset <- read_excel('ANX_2026_daner_fullANX_v12_woUTAH_11022026.gz.p4.clump.areator.sorted.1mhc (1).xls')
ANX_dataset
# Anxiety Dataset: 1320 x 23

```
Step 2: Get rid of unnecessary (for our purposes) columns
```{r}
ANX_dataset <- ANX_dataset[, -c(11:23)]
ANX_dataset
```


Step 3: Subset the data by adjusting and filtering p-values
First, Benjamin-Hochberg
```{r}
ANX_dataset$BH_P <- p.adjust(ANX_dataset$P, "BH")
#choose threshhold of 0.05
ANX_BH_adjusted = ANX_dataset[ANX_dataset$BH_P < 0.05, ] 
ANX_BH_adjusted
# ANX_BH_adjusted: 1320 x 24 ==> All p-values are significant with a 5% FDR

```
Second, Bonferroni
```{r}
ANX_dataset$Bon_P <- p.adjust(ANX_dataset$P, "bonferroni")
#choose threshhold of 0.05
ANX_Bon_adjusted = ANX_dataset[ANX_dataset$Bon_P < 0.05, ] 
ANX_Bon_adjusted
```
Step 4: Analyze the Odds Ratio
Odds Ratio is a score of whether a SNP has 
  higher odds (OR > 1), 
  lower odds (OR < 1), 
  or no association (OR = 1) 
of the outcome being ANX

Comparing ORs in relation to p-values helps us understand the significance of the ORs

```{r}
# Install ggplot2 if you don't have it
# install.packages("ggplot2")
library(ggplot2)

# Step 1: Transform the the OR and PVals
# Taking the log of both OR and PVals help the visualization to be more clear and discernable
ANX_dataset$logP <- -log10(ANX_dataset$P)
ANX_dataset$logOR <- log2(ANX_dataset$OR)

# Step 2: Calculate the BH/Bon thresholds (converted to -log10 scale for the Y-axis)
bh_thresh <- -log10(max(ANX_dataset$P[ANX_dataset$BH_P < 0.05])) 
      # Look at all the raw PVals who have the BH PVals less than 0.05
      # The largest of those PVals will be the "worst"-most PVal that passes the BH test
bonf_thresh <- -log10(0.05 / 1320) 
      # This is simply the Bonferonni Correction equation

# Step3: Create a column to color code the 2 tiers 
ANX_dataset$Tier <- "Benjamin-Hochberg (Pink)" 
      # All rows PVals labeled as BH right now
ANX_dataset$Tier[ANX_dataset$P < (0.05 / 1320)] <- "Bonferroni (Purple)" 
      # Now, only PVals under the Bonferroni Threshold are labeled 

# Steo 4: Plot the Volcano
ggplot(ANX_dataset, aes(x = logOR, y = logP, color = Tier)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_hline(yintercept = bh_thresh, linetype = "dashed") + # Lower line
  geom_hline(yintercept = bonf_thresh, linetype = "solid") + # Upper line
  geom_vline(xintercept = 0, linetype = "solid", color = "black") + # OR divide line
  theme_minimal(base_family = "serif") +
  scale_color_manual(values = c(
    "Benjamin-Hochberg (Pink)" = "pink",
    "Bonferroni (Purple)" = "#925EFF"
    )) +
  labs(title = "ANX Volcano Plot",
       x = "Log2(Odds Ratio)",
       y = "-Log10(P-value)",
       subtitle = "Visualizes the protective and risk factors of each SNPs in the GWAS study."
       )
    
```
This Volcano plot visualize the risk and protective factors of each SNPs in the GWAS studies. The rightmost SNPs, have OR > 1, meaning they have a higher probability of being associated with ANX. The uppermost SNPs, have the lowest p-values, and thus, their OR scores are most signifcant. From this we can deduce, the gene variants most associated with ANX. 


From this plot, we reaffirm that all the SNPs used in this GWAS studied were properly adjusted, as all p-values pass the Benjamin-Hochberg test. In addition, the more we can see a majority of the SNPs also have a significant p-value when adjusted with the Bonferroni correction as well. From the Volcano-plot, we are able to visualize THE MOST significant SNPs. We can cross-references these SNPs with those discovered as most significantly associated with ANX in the study. 

```{r}
install.packages("ggrepel")
library(ggrepel)
# Step 1: Create a new column called 'Label'
ANX_dataset$Label <- NA

# Step 2: Chose labeling threshold
# We choose to label dots above y-axis = 10 
# -log10(PVal) >= 11 ==> PVal < 1e-11
label_filter <- ANX_dataset$P < 1e-11
  # Column of Trues and Falses now

# Step 3: Copy the SNP names into the blank column ONLY for the rows that passed the filter
ANX_dataset$Label[label_filter] <- ANX_dataset$SNP[label_filter]
```

```{r}
ggplot(ANX_dataset, aes(x = logOR, y = logP, color = Tier)) +
  geom_point(alpha = 0.8, size = 2) +
  
  geom_text_repel(aes(label = Label), 
                  color = "black",       # Keeps the text black so it's easy to read
                  box.padding = 0.5,     # Gives the text bubbles some breathing room
                  max.overlaps = Inf,    # Forces R to draw every label we asked for
                  show.legend = FALSE) + # Stops the labels from messing up your legend

  geom_hline(yintercept = bh_thresh, linetype = "dashed") + # Lower line
  geom_hline(yintercept = bonf_thresh, linetype = "solid") + # Upper line
  geom_vline(xintercept = 0, linetype = "solid", color = "black") + # OR divide line
  theme_minimal(base_family = "serif") +
  scale_color_manual(values = c(
    "Benjamin-Hochberg (Pink)" = "pink",
    "Bonferroni (Purple)" = "#925EFF"
    )) +
  labs(title = "ANX Volcano Plot",
       x = "Log2(Odds Ratio)",
       y = "-Log10(P-value)",
       subtitle = "Visualizes the protective and risk factors of each SNPs in the GWAS study."
       )
```
Interesting further study ==> are the labeled SNPs physically close together on certain chromosomes?

```{r}
sig_SNPs <- ANX_dataset$Label
sig_SNPs <- sig_SNPs[!is.na(sig_SNPs)]
sig_SNPs

```

Odds Ratio: A score for odds of association with anxiety. Greater than 1 means strong odds, less than 1 is low odds, 1 is no association (N/A)
--> scores have been normalized with log2
P-value: Significance levels, measuring the validity of these odds values