---
title: "graph1"
output: html_document
---

```{r}
print("Hello, world!")
```

Step 1: Load Anxiety GWAS dataset
```{r}
install.packages("readxl")
library('readxl')

ANX_dataset <- read_excel('/Users/soumyakalle/Desktop/LUNE/Csbio10_project/31389910/ANX_2026_Daner.xls')
ANX_dataset
# Anxiety Dataset: 1320 x 23

```
Step 2: <b> Get rid of unnecessary (for our purposes) columns </b>
```{r}
ANX_dataset <- ANX_dataset[, -c(11:23)]
ANX_dataset
```


Step 3: Subset the data by adjusting and filtering p-values
First, Benjamin-Hochberg
```{r}
ANX_dataset$BH_P <- p.adjust(ANX_dataset$P, "BH")
#choose threshhold of 0.05
ANX_BH_adjusted = ANX_dataset[ANX_dataset$BH_P < 0.05, ] 
ANX_BH_adjusted
# ANX_BH_adjusted: 1320 x 24 ==> All p-values are significant with a 5% FDR

```
Second, Bonferroni
```{r}
ANX_dataset$Bon_P <- p.adjust(ANX_dataset$P, "bonferroni")
#choose threshhold of 0.05
ANX_Bon_adjusted = ANX_dataset[ANX_dataset$Bon_P < 0.05, ] 
ANX_Bon_adjusted
```
Step 4: Analyze the Odds Ratio
Odds Ratio is a score of whether a SNP has 
  higher odds (OR > 1), 
  lower odds (OR < 1), 
  or no association (OR = 1) 
of the outcome being ANX

Comparing ORs in relation to p-values help us understand the significance of the ORs

```{r}
# Install ggplot2 if you don't have it
# install.packages("ggplot2")
library(ggplot2)

# 1. Prepare the data: Calculate -log10(P) and log2(OR)
# We use log transformations to make the plot symmetrical and readable
ANX_dataset$logP <- -log10(ANX_dataset$P)
ANX_dataset$logOR <- log2(ANX_dataset$OR)

# 2. Define your two thresholds (converted to -log10 scale for the Y-axis)
bh_thresh <- -log10(max(ANX_dataset$P[ANX_dataset$BH_P < 0.05])) 
bonf_thresh <- -log10(0.05 / 1320)

# 3. Create a column to categorize your tiers for coloring
ANX_dataset$Tier <- "Silver (BH Only)"
ANX_dataset$Tier[ANX_dataset$P < (0.05 / 1320)] <- "Gold (Bonferroni)"

# 4. Plot the Volcano
ggplot(ANX_dataset, aes(x = logOR, y = logP, color = Tier)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_hline(yintercept = bh_thresh, linetype = "dashed") + # Lower line
  geom_hline(yintercept = bonf_thresh, linetype = "solid") + # Upper line
  geom_vline(xintercept = 0, linetype = "solid", color = "black") + # Neutral OR line
  theme_minimal() +
  labs(title = "Volcano Plot of Anxiety Loci",
       x = "Log2(Odds Ratio)",
       y = "-Log10(P-value)")
```


