---
title: "Project"
output: html_document
date: "2026-05-22"
---

```{r}
library('readxl')
Xanax <- read_excel('ANX_2026_daner_fullANX_v12_woUTAH_11022026.gz.p4.clump.areator.sorted.1mhc (1).xls')
install.packages("tidyverse")
library(ggplot2)
library(tidyverse)
Xanax %>%
  select(SNP, CHR, BP, P) %>% 
  # Convert chromosome number to a factor so each chromosome gets a distinct color
  mutate(CHR = as.factor(CHR)) %>% 
  arrange(CHR,BP) %>% 
  #chromosome numbers have been multiplied by 100 million to separate them visually on the graph
  mutate(Chromosome_Position = as.numeric(CHR)*1000000000+BP) %>% 
  
  #y intercept denotes a line of significance; from the article we pulled the data from, the line was
  ggplot(aes(Chromosome_Position, -log10(P), color= CHR))+
  geom_point(alpha=0.5) + geom_hline(yintercept = 7.30102999566, color = 'darkblue', size = 0.5, alpha = 0.5)
  #p-value of 5^(10^(-8)) or lower

```

```{r}
Xanax %>%
  select(SNP, CHR, BP, P) %>% 
  mutate(CHR = as.factor(CHR)) %>% 
  arrange(CHR,BP) %>% 
  # plots split into one panel per chromosome for easier per-chromosome reading
  ggplot(aes(BP, -log10(P), color= CHR))+
  geom_point(alpha=0.5)+
  facet_wrap(~CHR)
```

```{r}
for (i in 1:6) {
  start_chr <- (i-1)*5+1
  end_chr <- i*5
  if(end_chr>22) {
    end_chr<-22
  }
chr_batch <- as.numeric(start_chr:end_chr)
filt_xanax <- Xanax %>%
  select(SNP, CHR, BP, P) %>% 
  filter(as.numeric(CHR) %in% chr_batch)

if (nrow(filt_xanax)>0) {
  fx <- ggplot(filt_xanax, aes(BP, -log10(P), color=as.factor(CHR)))+
    geom_point(alpha=0.5)+
    facet_wrap(~CHR, nrow = 1, scales="free_x")+
    labs(title=paste("Chromosomes", start_chr, "to", end_chr))
  print(fx)
  }
}
```

```{r}
for (i in 1:5) {
  start_chr <- (i-1)*5+1
  end_chr <- i*5
  chr_batch <- as.numeric(start_chr:end_chr)
  filt_xanax <- Xanax %>%
    select(SNP, CHR, BP, P) %>% 
    filter(as.numeric(CHR) %in% chr_batch)
  if (nrow(filt_xanax)>0) {
    fx <- ggplot(filt_xanax, aes(BP, -log10(P), color=as.factor(CHR)))+
      geom_point(alpha=0.5)+
      lims(y=c(4,15))+
      facet_wrap(~CHR, nrow = 1, scales="free_x")+
      labs(title=paste("Chromosomes", start_chr, "to", end_chr))
    print(fx)
  }
}
```
<<<<<<< HEAD
Xanax <- read.csv('/Users/julia/Desktop/C and S Bio/ANX_2026_daner_fullANX_v12_woUTAH_11022026.gz.p4.clump.areator.sorted.1mhc.xls - daner_fullANX_v12_woUTAH_11022.csv')


```{r}
library(ggplot2)
library(tidyverse)
chr1_hits <- Xanax %>% 
  filter(CHR == 1, -log10(P) > 4)

# 2. Plot them exactly on a single line (y = 1)
ggplot(chr1_hits, aes(x = BP, y = 1)) +
  geom_point(alpha = 0.5, color = "skyblue", size = 3) +
  
  # This section strips away all the empty vertical y-axis space
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    
    # THIS IS THE MAGIC: Forces the plot to be short and wide like a literal line
    aspect.ratio = 0.05 
  ) +
  labs(title = "Chromosome 1: Base Pair Positions", x = "Base Pair Position (BP)")
mean(diff(sort(chr1_hits$BP)))
chr1_positions <- Xanax %>% 
  filter(CHR == 1) %>% 
  select(BP) %>% 
  arrange(BP)
print(chr1_positions)
```

